### 2022-04-24 greta/hopf_mcmc.Rmd
set test_t_max

```{r}
combined <-  compare_forecast(draws, train, test, simulate, vars = c("H", "P"))

@@

```{r}
combined <- compare_forecast(draws2, train, test, simulate, vars = c("H", "P"))

>>>>

```{r}
combined <-  compare_forecast(draws, train, test, simulate, vars = c("H", "P"),
                              test_reps, test_t_max)

@@

```{r}
combined <- compare_forecast(draws2, train, test, simulate, vars = c("H", "P"),
                              test_reps, test_t_max)

### 2022-04-24 greta/sn_mcmc.Rmd
set test_t_max

```{r}
library(tidyverse)
library(greta)
library(bayesplot)

set.seed(4242)
train_reps <- 100
train_t_max <- 250
test_t_max <- 250
test_reps <- 100

np.clip <- function(x, a, b) {
  if(x < a) return(a)
  if(x > b) return(b)
  x
}
```

```{r}
# View the bistable curve: 
#curve( (0.75 - x)**2 * (0.25 - x), 0.749999, 0.750001)
#curve(0*x,0,1,add = TRUE)

#curve((0.25 - x)**2 * (0.75 - x), 0, 1)
#curve(0*x,0,1,add = TRUE)
```

```{r}
step <- function(N, eta,  a = 0.75, b = 0.25) {
  N <- N + (a - N)**2 * (b - N) + eta
  N <- np.clip(N, 0, 1)
  N
}

# simulate
simulate <- function(N_init = 0.75,
                     t_max = 1250L,
                       p = list(
                       sigma = 1e-2,
                       a = 0.75, 
                       b = 0.25)) {
  eta <- rnorm(t_max, 0, p$sigma) # mu = 0, no drift
  N   <- numeric(t_max)
  N[1] <- N_init
  for (t in 1:(t_max-1)) {
    N[t+1] <- step(N[t], eta[t], a = p$a, b = p$b)
  }
  tibble::tibble(t = 1:t_max, N = N)
}

train <- purrr::map_dfr(1:train_reps, \(i) simulate(t_max=train_t_max), .id = "i")
test <- purrr::map_dfr(1:test_reps, \(i) simulate(t_max=test_t_max), .id = "i")

## show training data
train |> ggplot(aes(t, N, group=i)) + geom_line(alpha=0.2)
```

```{r}
gsims <- train |> group_by(i) |> mutate(xt1 = lead(N)) |> filter(t<max(t))

x_t <- gsims$N
x_t1 <- gsims$xt1

library(greta)
a <- uniform(0, 10)
b <- uniform(0, 10)
sigma <- uniform(0, 10)
mean <-  x_t + (a - x_t) ^ 2 * (b - x_t)
distribution(x_t1) <- normal(mean, sigma )
m <- model(a, b, sigma)
```

```{r}
mmcmc <- memoise::memoise(mcmc, cache = memoise::cache_filesystem("mcmc_cache"))

bench::bench_time({                 
  draws <- mmcmc(m, n_samples = 10000, warmup = 5000, chains = 4, verbose = FALSE)
})
```

```{r}
bayesplot::mcmc_trace(draws)
```

```{r}
## draw test_reps number of samples

posterior_samples <- 
  bind_rows(map(draws, as_tibble)) %>% 
  sample_n(test_reps)
posterior_sims <- posterior_samples %>%
  purrr::transpose() %>%
  map_dfr(function(q) simulate(t_max = test_t_max, p = q) ,.id = "i")

bind_rows(
  mutate(test, model="true"), 
  mutate(posterior_sims, model="predicted")
)|> 
  ggplot(aes(t, N, col=model, group=interaction(model,i))) +
           geom_line(alpha=0.1)

@@

```{r}
true <- tibble(a = 0.75, b = 0.25, sigma = 1e-2) |> gather(variable, value)
bind_rows(map(draws, as_tibble)) |>
  gather(variable, value) |> ggplot() + 
  geom_histogram(aes(value), bins = 30)  +
  geom_vline(data = true, aes(xintercept = value), col = "red", lwd = 1) + 
  facet_wrap(~variable, scales = "free")
```

# Scoring

First, a single 'observed' sample from the true:

```{r}
library(scoringRules)

scores <- function(observed, dat) {
  logsscore <- scoringRules::logs_sample(observed, dat)
  crpsscore <- scoringRules::crps_sample(observed, dat)
  data.frame(logs = mean(logsscore[-1]), crps =  mean(crpsscore[-1]))

}
# ensemble predictions
dat <- 
  posterior_sims |> 
  pivot_wider(id_cols = "t", names_from="i", values_from = "N") |> 
  select(-t) |> as.matrix()

```
```{r}
# score over all replicates:
bench::bench_time({
rep_scores <- 
  test |> 
  group_by(i) |> 
  group_map(~ scores(.x$N, dat)) |> 
  bind_rows()
})
```

```{r}
rep_scores |> summarise(across(.fns= base::mean))

library(patchwork)
ggplot(rep_scores) + geom_histogram(aes(crps)) + 
(ggplot(rep_scores) + geom_histogram(aes(logs)) )

>>>>

```{r setup, message=FALSE}
library(tidyverse, quietly = TRUE)
library(greta, quietly = TRUE)
library(bayesplot, quietly = TRUE)
source("R/utils.R")
source("R/hopf.R")
```

```{r}
set.seed(4242)
train_reps <- 1
train_t_max <- 250
test_t_max <- 250
test_reps <- 100
simulate <- simulate_sn
```

```{r}
pars = list(r = 1,
            K = 1,
            s = 0.1,
            h_init = .15,
            alpha = 0.0015,
            mu = 0,
            sigma = 0.02,
            t_init = 0
          )
t_max <- train_t_max + test_t_max

sim <- purrr::map_dfr(1:test_reps, 
                      \(i) simulate(t_max=t_max, p = p),
                      .id = "i")
train <- sim |> filter(t <= train_t_max, i <= train_reps)
test <- sim |> filter(t > train_t_max)
```

```{r}
m <- greta_model_stoch(train)
bench::bench_time({                 
  draws <- mmcmc(m, n_samples = 60000, warmup = 50000,
                 chains = 4, verbose = FALSE)
})
```

```{r}
## draw test_reps number of samples
combined <- compare_forecast(draws, train, test, simulate, vars = "N",
                              test_reps, test_t_max) 
write_csv(combined, "data/stochastic.csv.gz")
```

```{r}
scores <-
  rep_scores(combined, "N") |> 
  mutate(scenario="stochastic", 
         model="MCMC", 
         reps = train_reps)

write_csv(scores, "data/scores_stochastic.csv.gz")

@@

```{r}
bayesplot::mcmc_trace(draws)
plot_posteriors(draws, p)
```

```{r}
combined |> 
  group_by(t,type,variable) |> 
  summarise(mean = mean(value), sd = sd(value), .groups = "drop") |> 
 ggplot(aes(t, col=type)) + 
  geom_ribbon(aes(ymin = mean-2*sd, ymax = mean+2*sd, fill=type), alpha=0.5) +
  geom_line(aes(y=mean)) +
  geom_vline(aes(xintercept = train_t_max)) + facet_wrap(~variable, ncol=1)
```

```{r}
single <- 
  combined |> 
  filter(i %in% 1:train_reps) |> 
  ggplot(aes(t, value, col=type, group=interaction(i,type))) + 
    geom_line() +
    geom_vline(aes(xintercept = train_t_max)) + facet_wrap(~variable, ncol=1)
```

### 2022-04-24 greta/stochastic_mcmc.Rmd
set test_t_max

```{r}
## draw test_reps number of samples
combined <- compare_forecast(draws, train, test, simulate, vars = "N") 
write_csv(combined, "data/stochastic.csv.gz")
```

>>>>

```{r}
## draw test_reps number of samples
combined <- compare_forecast(draws, train, test, simulate, vars = "N",
                              test_reps, test_t_max) 
write_csv(combined, "data/stochastic.csv.gz")
```
